---
title: "Covid91 vaccine study Final2021"
author: "Shauna Breazeal"
date: "13 April 2021"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Methods
The first step is to isolate the subpopulations of Males, Females, Drug Users, and LGBTQ. 
```{r}
Maledata <- filter(FinalData,sex== "guy")
Maledata$sex <- as.factor(Maledata$sex)
```


# Males

## Barchart

```{r}
Males <- Maledata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This barchart shows that out of the approximately 340 Males who were infected with Covid-19, 57% were given the drug, and 43% were given the placebo.  

### Numerical Summaries  

```{r}
table1 <- xtabs(~treatment + infected, data=Maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```

The table shows that about 340 Males were infected with Covid-19 out of approximately 10,400 Males in the study. The table also shows that 3.72% of Males who received the drug contracted Covid-19 and 2.55% of Males who received the placebo contracted Covid-19. Lastly the table shows that Males who got the vaccine were about 50% more likely to contract Covid-19, this shows that the vaccine is not effective for Males.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value is so small at 0.0006181 that it is highly unlikely that the results were by chance. 
$H_0$ the null hypothesis is that the vaccine makes no difference in the Male population when it comes to contracting Covid-19. 
$H_0$ the alternative hypothesis is that the vaccine in fact does make a difference in whether or not a Male contracts Covid-19. 
Therefore, we reject the null hypothesis. 


# Females


# Drug Users


# LGBTQ


# Discussion and Conclusion


